Mechanistic Bayesian Forecasts of COVID19

### 2020-07-28 paper.tex
i hate git

% \begin{itemize}
% \item We choose tight priors based on literature
%  \item Fundamental unidentifiability due to renewal equation expression where $I(t)$ is a convolution between $R_t$ and Seiral interval
%  \item Cannot estimate both simultaneously 
%  \item Most flexibility comes in with beta and detection random walk
%  \item Ideal for forecasting, maybe not for inference
%  \end{itemize}
%  
  \subsection{Fitting}
We use the hamiltonian monte carlo algorithm implemented in numpyro (some citation here) to fit the model to data. That is, given a time series of confirmed cases ($\text{Cases}_{1:t,j}$) and confirmed deaths ($\text{Deaths}_{1:t,j}$) we use Bayesian inference (via HMC) to obtain

$$P(\bm{\theta} | \text{Cases}_{1:t,j}, \text{Deaths}_{1:t,j}) = \frac{P(\text{Cases}_{1:t,j},\text{Deaths}_{1:t,j}|\bm{\theta}) P(\bm{\theta})}{\int_{\bm{\theta}} P(\text{Cases}_{1:t,j}, \text{Deaths}_{1:t,j} |\bm{\theta})P(\bm{\theta})}}$$

@@

$$\bm{\theta} = \begin{pmatrix} 
\beta_{t,j} \\
\sigma_j \\
\gamma_j \\
\rho_j \\
\lambda_j \\
p_j \\
p_{d_j} \\ 
\text{case detection probability}_{t,j} \\
\text{death detection probability}_{j} \\
\text{case over-dispersion}_{j} \\
\text{death over-dispersion}_{j} 
\end{pmatrix}

%
% \begin{itemize}
% \item Fully Bayesian HMC on parameters
% \item Why we choose deterministic compartmental model with only uncertainty on parameters and observations
% \item Estimation in numpyro very fast
%   \end{itemize}
%   
%   
%   
 \section{Experimental Setup}

 \subsection{COVID-HUB}
 \begin{itemize}
 \item Real-time forecasting evaluation for 14 weeks starting April 20th 2020. 
 \item Forecasts submitted every Monday using incident data up until Sunday
 \item Cumulative forecasts generated by aggregation of incident forecasts
 \item 1-4 week ahead targets generated from 28 day ahead predictions 
 \end{itemize}

>>>>

% \begin{itemize}
% \item We choose tight priors based on literature
%  \item Fundamental unidentifiability due to renewal equation expression where $I(t)$ is a convolution between $R_t$ and Seiral interval
%  \item Cannot estimate both simultaneously 
%  \item Most flexibility comes in with beta and detection random walk
%  \item Ideal for forecasting, maybe not for inference
%  \end{itemize}
%  
  \subsection{Fitting}
We use the hamiltonian monte carlo algorithm implemented in numpyro (some citation here) to fit the model to data. That is, given a time series of confirmed cases ($C_{1:t}$) and confirmed deaths ($D_{1:t}$) we use Bayesian inference (via HMC) to obtain

$$f(\bm{\theta} | C_{1:t},D_{1:t}) \propto f(C_{1:t},D_{1:t} | \bm{\theta})f(\bm{\theta})$$

@@

$$\bm{\theta} =[\beta_{t} ,
\sigma ,
\gamma ,
\rho ,
\lambda ,
p_{c,t} ,
p_{d} ,
\sigma_c^2 ,
\sigma_d^2 ,
I_0 ,
E_0 ,
D_{1_0} ,
D_{2_0} ,
R_0 ]$$

We use the numpyro probabilistic framework to fit the model.

%
% \begin{itemize}
% \item Fully Bayesian HMC on parameters
% \item Why we choose deterministic compartmental model with only uncertainty on parameters and observations
% \item Estimation in numpyro very fast
%   \end{itemize}
%   
%   
%   
 \section{Experimental Setup}
 To evaluate our model, we examine two different scenarios. First, we describe the submission process and infrastructure used for the real time evaluation as part of the COVID-HUB consortium. Second, we describe the internal evaluation used to demonstrate our model enhancements improve accuracy over a naive compartmental model.

 \subsection{COVID-HUB}
 The COVID-HUB began soliciting forecasts in the beginning of April 2020 for 1-4 week ahead cumulative deaths. We began submitting the first version on April 20th 2020 and have since submitted forecasts every Monday from then until August 1st 2020. The forecasts use daily data up to and including the Sunday before submission the next Monday. The one week ahead forecast corresponds to the following Saturday, the two week ahead to the second following Saturday and so on. Our model went through three distinct iterations as we evaluated performance in real-time.

 \begin{itemize}

 \item \textbf{Version 1: April 20th 2020 through May 10th 2020}. Model had normal observation noise (instead of negative binomial) and a non-time varying case detection probability. Model was fit to cumulative deaths. 
  \item \textbf{Version 2: May 10th 2020 through May 24th 2020}. Model had normal observation noise (instead of negative binomial) and time varying case detection probability. Model was fit to cumulative deaths. 
    \item \textbf{Version 2: May 24th 2020 through August 1st 2020}. Model had negative binomial observation noise  and time varying case detection probability. Model was fit to incident deaths. 
 \end{itemize}

 These three versions highlight the complexities of forecasting during a real-time pandemic. 
% \begin{itemize}
% \item Real-time forecasting evaluation for 14 weeks starting April 20th 2020. 
% \item Forecasts submitted every Monday using incident data up until Sunday
% \item Cumulative forecasts generated by aggregation of incident forecasts
% \item 1-4 week ahead targets generated from 28 day ahead predictions 
% \end{itemize}
